\thetitle

### 2021-05-18 writing/discussion.tex
editng discussion about reticulated

At first glance, these measurements seem to contradict the results
of~\citet{vss-popl-2017}. They report an average slowdown of 2.5x and a
worst-case of 5.4x on the fully-typed Python benchmarks in Reticulated Python.
Unfortunately, the high cost of Shallow Racket seems closer to the truth.
There are at least four broad issues that skew the earlier results:
\begin{enumerate}

@@

\item The types of almost all local variables in the "fully typed" benchmarks is
{\tt Dynamic}, often due the impoverished nature of the type system, which does
not allow the ascription of precise types.  Code that ends up with the type {\tt
Dynamic} has fewer constraints to check at run-time---and much less information
to track in the blame map.

@@

\item On non-trivial programs, Reticulated Python suffers from the same high
overhead as Shallow Racket.  It runs the simplest benchmark from the GTP suite (translated into Python)
in approximately 40 seconds without blame and times out after 10 minutes when
the blame map is enabled.

\end{enumerate}
More work on Transient blame is needed to make an informed decision.

>>>>

At first glance, these measurements seem to contradict the results
of~\citet{vss-popl-2017}. They report an average slowdown of 6.2x and a
worst-case of 17.2x on the fully-typed Python benchmarks in Reticulated
Python with the blame map enabled.  Unfortunately, the orders-of-magnitude
higher cost of Shallow Racket seems closer to the truth.  There are at
least four broad issues that skew the earlier results:
\begin{enumerate}

@@

\item While Reticulated attempts to infer types for local variables,
  often local variables end up with type
{\tt Dynamic} due the impoverished nature of the type system,  which does
    not allow the ascription of precise
    types~\citet{g-dissertation-2020}. Code with the type {\tt
Dynamic} has fewer constraints to check at run-time---and much less information
to track in the blame map.

@@

\item Reticulated Python does not seem to employ any crucial
  implementation techniques that can mitigate the significant slowdown due to
    the blame map observed in Shallow Racket. For example Reticulated Python  runs the simplest
    benchmark from the GTP suite (translated into Python) in approximately
    40 seconds without blame but times out after 10 minutes with blame.

\end{enumerate}
More work on Transient blame is needed to make an informed decision about
its prospects as a viable production-level approach.
